---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Loading and preprocessing the data

```{r}
setwd("~/Documents/coursera/PA1_template")
data <- read.csv("./activity.csv", na.strings = "NA")
```

Transform the column date into a date.

```{r}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$steps <- as.numeric(data$steps)
data$interval <- as.numeric(data$interval)
```

##What is the mean total number of steps taken per day?
using the plyr package and the ggplot2 package

```{r}
library(plyr)
z <- ddply(data, ~date, summarise, sum=sum(steps))
library(ggplot2)
ggplot(z, aes(x=date,y=sum)) + 
        geom_histogram(binwidth=.5, colour="black", fill="white") + 
        geom_bar(stat="identity") +
        xlab("date") +
        ylab("total number of steps") +
        ggtitle("total numbers of steps per day")
```

report the mean and median of the total number of steps taken per day
```{r}
x <- ddply(data, ~date, summarise, meanSteps=mean(steps), medianSteps=median(steps))
print(x)
```

## What is the average daily activity pattern?
Plotting a 5-minute interval steps vs the averaged number of steps taken across all days.

```{r}
    plot(data$date, data$steps, ylab="steps", xlab="date", type = "l")
        lines(x$date, x$meanSteps, col = "red")
        legend("topright", lty = 1:1 , col = c("black","red"), legend = c("5-minute interval","average per day"))
```

Find out which interval is containing on average the maximun number of steps

```{r}
DataMeanInterval <- ddply(data, ~interval, summarise, meanSteps=mean(steps, na.rm = TRUE))
        maxSteps <- max(DataMeanInterval[,2])        
        q <- DataMeanInterval$meanSteps == maxSteps
        IntervalMaximalNumberOfSteps <- DataMeanInterval[q,1]
        IntervalMaximalNumberOfSteps <- IntervalMaximalNumberOfSteps[!is.na(IntervalMaximalNumberOfSteps)]
print(IntervalMaximalNumberOfSteps)
```

## Imputing missing values
To find the number of missing values in the dataset we will use the is.na function in combination with the sum function

```{r}
sum(is.na(data))
```

Create a new dataset where we're going to fill in the missing values of the dataset with the average of the mean of the 5 minutes interval
```{r}
dataNoNA <- data
n = 17568
for (i in 1:n) {
        if(is.na(dataNoNA[i,1]) == "TRUE") {
        z <- dataNoNA[i, 3]
        step <- DataMeanInterval$interval == z
        step <- DataMeanInterval[step,2]
        step <- step[!is.na(step)]
        dataNoNA[i,1] <- step
}
}

```

The histogram&mean and median are changing when there are no missing values present.

```{r}
z2 <- ddply(dataNoNA, ~date, summarise, sum=sum(steps))
ggplot(z2, aes(x=date,y=sum)) + 
        geom_histogram(binwidth=.5, colour="black", fill="white") + 
        geom_bar(stat="identity") +
        xlab("date") +
        ylab("total number of steps") +
        ggtitle("total numbers of steps per day")

x <- ddply(dataNoNA, ~date, summarise, meanSteps=mean(steps), medianSteps=median(steps))
print(x)
```

## Are there differences in activity patterns between weekdays and weekends?
create a newcolumn called wDay

``` {r}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
dataNoNA$wDay <-  factor((weekdays(dataNoNA$date) %in% weekdays1)+1L,
      levels=1:2, labels=c('weekend', 'weekday'))
```

panel plot

```{r}
z3 <- ddply(dataNoNA, .(interval,wDay), summarise, mean=mean(steps))
library(lattice)
xyplot(mean~interval | factor(wDay), data=z3, pch=19,
                xlab="interval",  ylab="Number of Steps",layout=c(1,2),type="l")
```